### 2013-04-25 inst/examples/BUGS/2013-04-19-bayesian-parametric-uncertainty-sdp.Rmd
updated examples

@@

  # compute mean with monte carlo samples
  mu <- sapply(indices, function(i) { # Cols are parameter samples
    p <- unname(pardist[i,c(4,1,5)])          # Only, we need to add the measurement and process noise  
    unname(sapply(x_grid, f, h, p))   # rows are possible values of x_t
  })
  mu <- data.frame(t(mu)) # Let's have cols as possible x_t, rows as parameters, and type as data.frame

  F_true <- 
    sapply(mu, function(cur_state){    # For each x_t

      bypar <- sapply(cur_state, function(m){ # For each parameter value
        if(snap_to_grid(m,x_grid) < x_grid[2]){ # 
          out <- numeric(length(x_grid))
          out[1] <- 1
          out
        } else {
          out <- dlnorm(x_grid/m, 0, sigma_g) # should be using the estimate process noise!
        }})  
      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
      ave_over_pars / sum(ave_over_pars)
      })

  F_true <- t(F_true)

@@

```{r}
pardist <- mcmcall
pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
pardist[,5] = p[3] + rnorm(100, 0, 0.000001)
```

>>>>

```{r libs, include=FALSE, cache=FALSE}
library(pdgControl)
library(nonparametricbayes)
library(reshape2)
library(ggplot2)
library(knitcitations)
```

```{r setupall, echo=FALSE}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE)
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

@@

  # Compute the matrix
  F_true <- 
    sapply(x_grid, function(x_t){           # For each x_t

      bypar <- sapply(indices, function(i){ # For each parameter value
        p <- unname(pardist[i,c(4,1,5)])    # parameters for the mean, at current sample 
        mu <- f(x_t,h,p)
        est_sigma_g <- pardist[i,2]         # Variance parameter

        if(snap_to_grid(mu,x_grid) < x_grid[2]){ # handle the degenerate case 
          out <- numeric(length(x_grid))
          out[1] <- 1
          out
        } else {
          out <- dlnorm(x_grid/mu, 0, est_sigma_g) 
        }
      })  
      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
      ave_over_pars / sum(ave_over_pars)
      })
  F_true <- t(F_true)

@@

```{r}
pardist <- mcmcall
pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
pardist[,2] = sigma_g + rnorm(100, 0, 0.000001)
pardist[,5] = p[3] + rnorm(100, 0, 0.000001)
```

### 2013-04-25 inst/examples/BUGS/allen.Rmd
updated examples

```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
f <- RickerAllee
p <- c(2, 10, 5) 
K <- 10
allee <- 5
```

@@

```{r obs, dependson="sdp-pars"}
obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=1, 
                harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
```

@@

@@

```{r}
jags.params=c("K","logr0","logtheta","iR","iQ"); # Don't need to save "x"
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]),"iQ"=1/0.05,"iR"=1/0.1,"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
time<-system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-GammaPrior.txt")
)         
time <- unname(time["elapsed"]);
```

@@

```{r}
theta <- exp(mcmcall[,"logtheta"])
theta_dist <- hist(theta, freq=FALSE)
theta_dist$mids
theta_dist$density
delta <- theta_dist$mids[2] - theta_dist$mids[1]
```

evaluating the value function given $f$ fixed at each of `theta_dist$mids` multiplied by  `theta_dist$density * delta` and summed over each $\theta$ is going to be slow for a single parameter but simply unrealistic for higher dimensions.  Instead we will rely on Monte Carlo sampling of parameter values from their posteriors,

```{r}
mc_n <- 100
thetas <- sample(theta, mc_n, replace=T)

>>>>

```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE)
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### Model and parameters

```{r}
f <- RickerAllee
p <- c(2, 10, 5) 
K <- p[2]
allee <- p[3]
```

@@

### Sample Data

```{r obs, dependson="sdp-pars"}
obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=15, 
                harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
```

## Maximum Likelihood

@@

## Non-parametric Bayes

```{r gp-priors}
#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
s2.p <- c(5,5)  
tau2.p <- c(5,1)
d.p = c(10, 1/0.1, 10, 1/0.1)
nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
beta0_prior <- function(x, tau) dnorm(x, 0, tau)
beta = c(0)
priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
```

```{r}
  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
          m0r1=FALSE, corr="exp", trace=TRUE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
```

```{r}
  posteriors_plot(gp, priors) # needs trace=TRUE!
```

#### Parametric Bayes

@@

```{r}
jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]),"iQ"=1/0.05,"iR"=1/0.1,"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
time<-system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-GammaPrior.txt")
)         
time <- unname(time["elapsed"]);
```

#### Convergence diagnostics for parametric bayes

@@

```{r}
MaxT = 1000
matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

pardist <- mcmcall
pardist[,4] = exp(pardist[,4])
pardist[,5] = exp(pardist[,5])

matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

OPT = list(gp_D = opt_gp$D, true_D = opt_true$D, est_D = opt_estimated$D, alt_D = opt_alt$D, par_bayes = opt_par_bayes$D)
```

```{r}
policies <- melt(data.frame(stock=x_grid, 
                            GP = x_grid[opt_gp$D], 
                            MLE = x_grid[opt_estimated$D],
                            True = x_grid[opt_true$D],
                            Ricker_MLE = x_grid[opt_alt$D],
                            Parametric.Bayes = x_grid[opt_par_bayes$D]),
                   id="stock")
names(policies) <- c("stock", "method", "value")
policy_plot <- ggplot(policies, aes(stock, stock - value, color=method)) +
  geom_line(lwd=1.2, alpha=0.8) + 
  xlab("stock size") + ylab("escapement")  +
  scale_colour_manual(values=cbPalette)
```

```{r}
sims <- lapply(OPT, function(D){
  set.seed(1)
  lapply(1:100, function(i) 
    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
  )
})

dat <- melt(sims, id=names(sims[[1]][[1]]))
dt <- data.table(dat)
setnames(dt, c("L1", "L2"), c("method", "reps")) 
setkey(dt, method) # change the ordering

ggplot(dt) + 
    geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))

ggplot(dt) + 
    geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))

### 2013-04-25 inst/examples/BUGS/par_uncertainty_sdp.Rmd
updated examples

```{r}
f_transition_matrix <- function(f, p, x_grid, h_grid, sigma_g, pardist){
  lapply(h_grid, par_F, f, p, x_grid, sigma_g, pardist)
}

par_F <- function(h, f, p, x_grid, sigma_g, pardist, n_mc = 100){

  # Set up monte carlo sampling 
  d <- dim(pardist)
  indices <- round(runif(n_mc,1, d[1]))

  # compute mean with monte carlo samples
  mu <- sapply(indices, function(i) { # Cols are parameter samples
    p <- unname(pardist[i,c(4,1,5)])          # Only, we need to add the measurement and process noise  
    unname(sapply(x_grid, f, h, p))   # rows are possible values of x_t
  })
  mu <- data.frame(t(mu)) # Let's have cols as possible x_t, rows as parameters, and type as data.frame

  F_true <- 
    sapply(mu, function(cur_state){    # For each x_t

      bypar <- sapply(cur_state, function(m){ # For each parameter value
        if(snap_to_grid(m,x_grid) < x_grid[2]){ # 
          out <- numeric(length(x_grid))
          out[1] <- 1
          out
        } else {
          out <- dlnorm(x_grid/m, 0, sigma_g) # should be using the estimate process noise!
        }})  
      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
      ave_over_pars / sum(ave_over_pars)
      })

  F_true <- t(F_true)

}

snap_to_grid <- function(x, grid) sapply(x, function(x) grid[which.min(abs(grid - x))])

# internal helper function
rownorm <- function(M) 
  t(apply(M, 1, function(x){ 
    if(sum(x)>0){
      x/sum(x)
    } else {
      out <- numeric(length(x))
      out[1] <- 1
      out
    }
  }))
```

```{r}
knit("allen.Rmd")

pardist <- mcmcall
#pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
#pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
#pardist[,5] = p[3] + rnorm(100, 0, 0.000001)

sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
#opt <- find_dp_optim(sdp, x_grid, h_grid, OptTime, xT, profit, delta, reward=0)

SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)

require(reshape2)
policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")

ggplot(policies, aes(stock, value, color=variable)) + geom_point(alpha=.5) + xlab("stock size") + ylab("harvest")

ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
```

>>>>

Run the Bayesian analysis to obtain posterior distributions for parameters.

```{r, echo=TRUE, include=FALSE}
knit("allen.Rmd")
```

```{r}
pardist <- mcmcall
```

Transform parameters back

```{r}
pardist[,4] = exp(pardist[,4])
pardist[,5] = exp(pardist[,5])
```

Compute optimal policy

```{r}
sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Compare to the case without parameter uncertainty (growth noise only)

```{r}
SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Plot results

```{r}
require(reshape2)
policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
```
